---
title: "Download WQ Data from winter 1990"
author: "Alejandra Botero-Acosta"
date: "10/14/2022"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: inline
---


Last updated OCt, 2022 by Alejandra Botero.    
**This script is used to download from the USGS Water Quality portal.**  
Final raw dataset downloaded saved as allHUCdataV2.RData

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_chunk$set(eval = FALSE)#include output without rerunning
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r, include=FALSE}
library(tidyverse) #data cleaning/loading
library(dataRetrieval) #USGS data import package
```

```{r, include=FALSE}
saveDataPath <- "/home/boteroacostaa/P_and_N_Alejandra_codes/data/downloaded_WQP_1990-2022/"
```


# Download data from WQP  
Save workspace the first time is run. (Not necessary to run again! import workspace or file by file (see next chunk))
```{r, eval = FALSE}

#huc 10 had to be divided in 4-digit huc to be downloaded
dataList <- list("05", "06", "07", "08", "1001;1002;1003;1004;1005;1006;1007", "1008;1009;1010;1011;1012;1013;1014;1015", "1016;1017;1018;1019;1020;1021;1022;1023", "1024;1025;1026;1027;1028;1029;1030", "11")

namesList <- list("05", "06", "07", "08", "10_01_07", "10_08_15", "10_16_23", "10_24_30", "11")

#define if downloading from start date
WQPstartdate <- "12-01-1989" 

i=1
for(huc in dataList){
  HUCdata<-readWQPdata(huc=huc, characteristicType="Nutrient", siteType="Stream", startDateLo=WQPstartdate)
  huc_name<-namesList[i]
  #test with no startdate
  #HUCdata<-readWQPdata(huc=huc, characteristicType="Nutrient", siteType="Stream")
  #save downloaded data to specified filepath
  write.csv(HUCdata, paste0(saveDataPath, "HUC", huc_name, "data.csv"), row.names = F)
  #assign data to huc##data variable in environment
  assign(paste0("HUC", huc_name, "data"), HUCdata)
  #save also R dataframe
  saveRDS(HUCdata,file=paste0(saveDataPath, "HUC", huc_name, "data.RData"))
  print(paste0("Data loaded into global environment and named as ", paste0("huc", huc, "data")))
  i=i+1
}
```

When importing Rdata file by file (alternative option to importing the whole workspace)
```{r, eval = FALSE}
namesList <- list("05", "06", "07", "08", "10_01_07", "10_08_15", "10_16_23", "10_24_30", "11")

for (huc_name in namesList){
  temporal<-readRDS(paste0(saveDataPath, "HUC", huc_name, "data.RData"))
  #assign data to huc##data variable in environment
  assign(paste0("HUC", huc_name, "data"), temporal)
}
```

Use following chunk when data types don't match try binding first (next step) to check for errors due to data type
```{r, eval = FALSE}
class(HUC05data$DetectionQuantitationLimitMeasure.MeasureValue) <- "double"
class(HUC06data$DetectionQuantitationLimitMeasure.MeasureValue) <- "double"
class(HUC07data$DetectionQuantitationLimitMeasure.MeasureValue) <- "double"
class(HUC08data$DetectionQuantitationLimitMeasure.MeasureValue) <- "double"
class(HUC11data$DetectionQuantitationLimitMeasure.MeasureValue) <- "double"
class(HUC10_08_15data$DetectionQuantitationLimitMeasure.MeasureValue) <- "double"
class(HUC10_16_23data$DetectionQuantitationLimitMeasure.MeasureValue) <- "double"
class(HUC10_24_30data$DetectionQuantitationLimitMeasure.MeasureValue) <- "double"
```

Bind dataframes to combine and save dataframe in memory (next time just import this dataframe)
```{r, eval = FALSE}
allHUCdata<-dplyr::bind_rows(HUC05data, HUC06data, HUC07data, HUC08data, HUC10_01_07data, HUC10_08_15data, HUC10_16_23data, HUC10_24_30data, HUC11data)
tot_num_rec<-nrow(allHUCdata)
#save R dataframe. COMMENTED TO AVOID OVERWRITING
#save(allHUCdata,file=paste0(saveDataPath, "allHUCdata.RData"))
#save(tot_num_rec,file=paste0(saveDataPath, "tot_num_rec.RData"))
```

Identify unique parameters in field: CharacteristicName
```{r, eval = FALSE}
#allHUCdata$ResultMeasure.MeasureUnitCode<-as.character(fullresults$ResultMeasure.MeasureUnitCode)
allParamterNAmes<-allHUCdata%>%distinct(CharacteristicName)
#save R dataframe. COMMENTED TO AVOID OVERWRITING
#save(allParamterNAmes,file=paste0(saveDataPath, "allParamterNAmes.RData"))
```

Resulting unique parameters in field: CharacteristicName
```{r, echo = FALSE} 
#ECHO=False hides code, shows result
#imported to show result when rendering html
load(paste0(saveDataPath, "allParamterNAmes.RData"))
allParamterNAmes
```

Resulting total number of records:
```{r, echo = FALSE} 
#ECHO=False hides code, shows result
#imported to show result when rendering html
load(paste0(saveDataPath, "tot_num_rec.RData"))
tot_num_rec
```

Identify dates in database
```{r, eval = FALSE} 
alldates<-as.Date(allHUCdata$ActivityStartDate, format = "%Y-%m-%d")
min_date<-min(alldates)
max_date<-max(alldates)

#save R dataframe. COMMENTED TO AVOID OVERWRITING
#save(min_date,file=paste0(saveDataPath, "min_date.RData"))
#save(max_date,file=paste0(saveDataPath, "max_date.RData"))

```

Resulting max an min dates:
```{r, echo = FALSE} 
#ECHO=False; runs, hides code, shows result
#imported to show result when rendering html
load(paste0(saveDataPath, "min_date.RData"))
load(paste0(saveDataPath, "max_date.RData"))
min_date
max_date
```
Filtering dates to remove anomalous ones, creating dataset V2
```{r, eval = FALSE} 
allHUCdataV2<-filter(allHUCdata, ActivityStartDate >=  "1989-12-01" & ActivityStartDate <= "2022-09-30")
tot_num_recV2<-nrow(allHUCdataV2)
#save R dataframe. COMMENTED TO AVOID OVERWRITING
#save(allHUCdataV2,file=paste0(saveDataPath, "allHUCdataV2.RData"))
#save(tot_num_recV2,file=paste0(saveDataPath, "tot_num_recV2.RData"))
```


Identify dates in database V2
```{r, eval = FALSE} 
alldatesV2<-as.Date(allHUCdataV2$ActivityStartDate, format = "%Y-%m-%d")
min_dateV2<-min(alldatesV2)
max_dateV2<-max(alldatesV2)

#save R dataframe. COMMENTED TO AVOID OVERWRITING
#save(min_dateV2,file=paste0(saveDataPath, "min_dateV2.RData"))
#save(max_dateV2,file=paste0(saveDataPath, "max_dateV2.RData"))

```


Resulting max an min dates afte removing anomalous dates (V2):
```{r, echo = FALSE} 
#ECHO=False; runs, hides code, shows result
#imported to show result when rendering html
load(paste0(saveDataPath, "min_dateV2.RData"))
load(paste0(saveDataPath, "max_dateV2.RData"))
min_dateV2
max_dateV2
```


Resulting total number of records after removing anomalous dates (V2):
```{r, echo = FALSE} 
#ECHO=False hides code, shows result
#imported to show result when rendering html
load(paste0(saveDataPath, "tot_num_recV2.RData"))
tot_num_recV2
```

```{r, echo = FALSE} 
#ECHO=False; runs, hides code, shows result

```

```{r, include=FALSE}
#include_false, runs, hides code, hides result
```

```{r, eval = FALSE} 
#eval=false, doesn't run
```
